Stellar Bars in Isolated Gas-Rich Galaxies Do Not Slow Down

### 2022-04-19 paper/ms.tex
Add more intro paragraphs

We have performed a simulation of a disk galaxy using the finite-volume
gravity-hydrodynamics code AREPO \cite{2010MNRAS.401..791S}. In AREPO, the
fluid is discretized as a moving Voronoi mesh. We use the additional physics
in the galaxy formation module Stars and MUltiphase Gas in GaLaxiEs (SMUGGLE)
\cite{2019MNRAS.489.4233M}. SMUGGLE is a comprehensive and self-consistent
galaxy formation model with a wide range of physical processes, including
radiative heating/cooling, star formation, and stellar feedback. More detailed
information on SMUGGLE is given in the Methods section.

Our disk galaxy is a modified version of the GALAKOS
model\cite{2020ApJ...890..117D}, which consists of a radially exponential and
vertically isothermal stellar disk along with a
Hernquist\cite{1990ApJ...356..359H} dark matter halo and bulge. After
$\sim2.5\,\textrm{Gyr}$ of evolution, the GALAKOS disk forms a bar consistent
with the Milky Way bar in terms of pattern speed and length
($\sim40\,\textrm{km}/\textrm{s}/\textrm{kpc}$ and $\sim4.5\,\textrm{kpc}$,
respectively). We modify this setup by introducing a gas phase after $\sim
1.5\,\textrm{Gyr}$ of evolution. This initial gas phase includes a hole in the
center $4\,\textrm{kpc}$ and a constant surface density profile out to
$9.3\,\textrm{kpc}$. When the gas is added to the simulation, the bar has
already formed. More details on this configuration are given in the Methods
section.

>>>>

It has long been known that non-axisymmetric features in a stellar disk act to
redistribute mass and angular momentum.\cite{1972MNRAS.157....1L} The
interaction between a bar and a spheroid (either a dark matter halo or stellar
bulge) has received considerable interest.\cite{1984MNRAS.209..729T,
1985MNRAS.213..451W} A reasonable stellar disk in isolation seems to always
generate a bar instability,\cite{1971ApJ...168..343H} but the presence of a
static dark matter halo acts to stabilize the disk against the bar
instability.\cite{1973ApJ...186..467O, 1976AJ.....81...30H} Later, a number of
numerical simulations have confirmed that a live dark matter halo generically
exerts a negative torque on the bar, causing its pattern speed to decrease and
its length and strength to increase\cite{1992ApJ...400...80H,
2000ApJ...543..704D, 2002MNRAS.330...35A, 2002ApJ...569L..83A,
2003MNRAS.341.1179A, 2003MNRAS.346..251O, 2005MNRAS.363..991H,
2006ApJ...637..214M, 2007MNRAS.375..460W, 2009ApJ...697..293D}. Intuitively,
the bar excites a wake of resonant material in the dark matter halo in a
manner similar to dynamical friction. This wake lags and thus exerts a
negative torque on the bar.

The pre-eminent expectation that the dark matter halo acts to slow the bar
down conflicts with observational estimates of bar pattern speeds. Bar
rotation rates are typically classified by the dimensionless ratio
\begin{equation}
\Rot = \RCR/R_b\textrm{,}
\end{equation}
where \RCR{} is the radius of corotation and $R_b$ is the length of the
bar.\footnote{The radius of corotation \RCR{} is defined for circular orbits as
the radius at which the orbital frequency is equal to the pattern speed,
$\Omega_p$, of a given non-axisymmetric feature. In a galaxy with a constant
circular velocity $V_c$, it is given by $\RCR = V_c / \Omega_p$.} Galaxies
with $\Rot < 1.4$ are considered ``fast rotators'' while galaxies with $\Rot >
1.4$ are considered ``slow rotators.''\cite{2000ApJ...543..704D} Galaxies with
$\Rot < 1$ are not thought to be stable.\cite{1980AA....81..198C}
Observational estimates of the pattern speeds of bars indicate that nearly all
galaxies have $1 < \Rot < 1.4$.\cite{2011MSAIS..18...23C, 2015AA...576A.102A,
2019MNRAS.482.1733G}

The role of gas on the evolution of the bar is less well-understood. Since the
gas phase typically only contributes about $10-20\%$ of the mass of a galaxy
at the present day, one might naively expect it to have a subdominant effect
on the bar. However, because gas is collisional, it can participate in
non-resonant angular momentum exchange with the bar. Thus, numerical work has
shown that the gas phase has an outsized influence on the
bar.\cite{2010ApJ...719.1470V, 2013MNRAS.429.1949A}

We have performed a simulation of a disk galaxy using the finite-volume
gravito-hydrodynamics code AREPO.\cite{2010MNRAS.401..791S} In AREPO, the
fluid is discretized as a moving Voronoi mesh. We use the additional physics
in the galaxy formation module Stars and MUltiphase Gas in GaLaxiEs (SMUGGLE)
\cite{2019MNRAS.489.4233M}. SMUGGLE is a comprehensive and self-consistent
galaxy formation model with a wide range of physical processes, including
radiative heating/cooling, star formation, and stellar feedback. More detailed
information on SMUGGLE is given in the Methods section. Our disk galaxy is a
modified version of the GALAKOS model\cite{2020ApJ...890..117D}, which
consists of a stellar disk, stellar bulge, and dark matter halo. After
$\sim2.5\,\textrm{Gyr}$ of evolution, the GALAKOS disk forms a bar consistent
with the Milky Way bar in terms of pattern speed and length
($\sim40\,\textrm{km}/\textrm{s}/\textrm{kpc}$ and $\sim4.5\,\textrm{kpc}$,
respectively). We modify this setup to include a gas phase, the details of
which is given in the Methods section.
